---
title: "Ridership Prediction - SVM Model"
author: "Section1-Team7"
date: "December 10, 2017"
output: html_document
---

## Run the code for SVM static, 2-week dynamic and 1-day dynamic  models.

# Assigning required libraries and working directory

```{r,warning=FALSE}
library(mlr)
library(ggplot2)
library(dplyr)
library(caret)
library(gridExtra)
library(lubridate)
library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyr)
library(data.table)
library(zoo)
setwd("D:/FT1 - Team 7/Data/Peiwen/Predictive Final Folder v1")
```

# Importing Ridership, weather, school information to complete major portion of data extraction

```{r,warning=FALSE}
usage_pred <- read.csv("Usage.csv")
usage_pred <- subset(usage_pred, select = c("Date","Ridership"))
#usage_pred$Date <- format(as.Date(usage_pred$Date), format = "%m/%d/%Y")
weather_pred <- read.csv("Weather.csv")
#weather_pred$Date <- format(as.Date(weather_pred$Date), format = "%m/%d/%Y")
school_pred <- read.csv("School.csv")
school_pred$Status <- toupper(school_pred$Status)
school_pred1 <- school_pred[,1:3]
school_pred_new <- reshape(school_pred1,timevar = "DistrictName",idvar = c("Date"),direction = "wide")
school_pred_new %>% mutate_if(is.factor,as.character) -> school_pred_new_char
date_list <- subset(usage_pred,select = c("Date"))
date_list %>% mutate_if(is.factor,as.character) -> date_list_char
school_pred_new2 <- merge(date_list_char,school_pred_new_char,by.x = "Date", by.y = "Date", all.x = TRUE)
school_pred_new2$`usage_pred$Date` <- as.Date(school_pred_new2$Date,"%m/%d/%Y")
school_pred_new2 %>% arrange(`usage_pred$Date`) -> school_pred_new2
```

# First, we treat the school information as it is not in a machine learning format. The school information is present only after 2014 August 21st. Hence, we filter out that information out.

```{r,warning=FALSE}
school_pred_new2 %>% filter(`usage_pred$Date` >= "2014-08-21") -> schoolfill1
schoolfill2 <- schoolfill1[,c(2:47)]
schoolfill2[schoolfill2 == "OFF"] <- NA 
schoolfill2$Date <- schoolfill1[,1]
schoolfill2 %>% na.locf(.,na.rm = FALSE) -> schoolfill3
schoolfill3[is.na(schoolfill3)] <- "END"
schoolfill3 %>% select(Date,everything()) -> schoolfill3
schoolfill1_1 <- schoolfill1[,c(2:47)]
schoolfill3_1 <- schoolfill3[,-1]
schoolfill3_1[schoolfill1_1 == "OFF"] <- "OFF"
schoolfill3_1$Date <- schoolfill2$Date
schoolfill3_1 %>% select(Date,everything()) -> schoolfill3_1
schoolfill3_1$Date <- as.Date(schoolfill3_1$Date,"%m/%d/%Y")
```

# We then treat the 2013 and 2014 months for which don't have school information by using the school information from 2014 August.

```{r,warning=FALSE}
calender_pred <- read.csv("updated_calendar_holidays.csv")
calender_pred %>% mutate_if(is.factor,as.character) -> calender_pred_char
calender_pred_char$Date <- as.Date(calender_pred_char$Date,"%m/%d/%Y")
calender_pred_char$LYDate <- as.Date(calender_pred_char$LYDate,"%m/%d/%Y")
calender_pred_char$X2YDate <- as.Date(calender_pred_char$X2YDate,"%m/%d/%Y")
calender_pred_char$LYHolidayDate <- as.Date(calender_pred_char$LYHolidayDate,"%m/%d/%Y")
calender_pred_char$X2YHolidayDate <- as.Date(calender_pred_char$X2YHolidayDate,"%m/%d/%Y")

schoolfill3_2 <- merge(schoolfill3_1,calender_pred_char,by = "Date", all.x = TRUE)

# Using 2014 information to fill for 2013 dataset

school_pred_new2 %>% filter(`usage_pred$Date` < "2014-08-21") -> schoolpre2014
schoolpre2014 <- schoolpre2014[,-1]
year2yeardates <- calender_pred_char[,c(1,3)]
schoolpre2014_1 <- merge(schoolpre2014,year2yeardates,by.x = "usage_pred$Date", by.y = "X2YDate", all.x = TRUE)
schoolpre2014_1 %>% select(Date,everything()) %>% select(c(Date)) -> schoolpre2014_2

schoolpre2014_3 <- merge(schoolpre2014_2,schoolfill3_2,by.x = "Date", all.x = TRUE)
schoolpre2014_3 %>% select(-c(LYDate,X2YDate,LYHolidayDate,X2YHolidayDate,Holiday)) -> schoolpre2014_3
schoolpre2014_4 <- merge(schoolpre2014_3,schoolpre2014_1[,c("Date","usage_pred$Date")],by = "Date", all.x = TRUE)
schoolpre2014_4 %>% select(`usage_pred$Date`,everything()) %>% select(-c(Date)) -> schoolpre2014_4
schoolpre2014_5 <- merge(schoolpre2014_4,calender_pred_char[,c("Date","Holiday")],by.x = "usage_pred$Date", by.y = "Date", all.x = TRUE) 
schoolpre2014_5[schoolpre2014_5$Holiday != "NULL",2:47] <- "OFF"
names(schoolpre2014_5)[names(schoolpre2014_5) == 'usage_pred$Date'] <- 'Date'

schoolfill3_2 %>% select(-c(LYDate,X2YDate,LYHolidayDate,X2YHolidayDate)) -> schoolfill3_2

# Combining
schoolinfofinal <- rbind(schoolpre2014_5,schoolfill3_2)
```

# We merge the ridership, weather and treated school information to form the first part of the dataset.

```{r,warning=FALSE}
schoolinfofinal %>% mutate_if(is.character,as.factor) -> school_pred
#school_pred$Date <- as.Date(school_pred$Date)
#weather_pred$Date <- as.Date(weather_pred$Date, format = "%m/%d/%Y")
usage_weather <- merge(usage_pred,weather_pred,by.x = "Date",by.y = "Date",all.x = TRUE)
#school_pred <- read.csv("School.csv")
usage_weather$Date <- mdy(usage_weather$Date)
usage_weather_school <- merge(usage_weather,school_pred,by.x = "Date",by.y = "Date",all.x = TRUE)

#usage_weather_school$Date <- mdy(usage_weather_school$Date)
usage_weather_school$day <- weekdays(as.Date(usage_weather_school$Date))
finaldataset_rwshd <- usage_weather_school
finaldataset_rwshd$day <- as.factor(finaldataset_rwshd$day)
```

# We removed Christmas information from the dataset because park is closed on Christmas. 

```{r,warning=FALSE}
finaldataset_rwshd <- finaldataset_rwshd[finaldataset_rwshd$Holiday != "Christmas Day",]
calender_pred$Date <- as.Date(calender_pred$Date, format = "%m/%d/%Y")
calender_pred$LYDate <- as.Date(calender_pred$LYDate, format = "%m/%d/%Y")
calender_pred$X2YDate <- as.Date(calender_pred$X2YDate, format = "%m/%d/%Y")
finaldataset_rwshd_lyr1 <- merge(finaldataset_rwshd,calender_pred[,c("Date","LYDate","X2YDate")],by.x = "Date",by.y = "Date",all.x = TRUE)

finaldataset_rwshd_lyr2 <- merge(finaldataset_rwshd_lyr1,finaldataset_rwshd[,c("Date","Ridership")],by.x = "LYDate",by.y = "Date",all.x = TRUE)


```

# 2012 ridership was provided separately. We add this ridership information to the dataset. For modeling purposes, we create 7-day lag, 1-day lag, 2 day lag, 1 year lag and 1 year + 1 day lag. We also observe that part was closed on thanksgiving 2016. Hence, we treat this field with previous year's mean values.

```{r,warning=FALSE}
# add rides in 2012
ridership2012<-read.csv("2012Ridership.csv")
ridership2012$Date <-as.Date(ridership2012$Date)
ridership2012 <- ridership2012%>%group_by(Date) %>%summarise(Ridership = sum(Total.Uses))

finaldataset_rwshd_lyr3 <- merge(finaldataset_rwshd_lyr2,ridership2012,by.x = "LYDate",by.y = "Date",all.x = TRUE)
finaldataset_rwshd_lyr3$Ridership.y <- ifelse(is.na(finaldataset_rwshd_lyr3$Ridership.y),finaldataset_rwshd_lyr3$Ridership,finaldataset_rwshd_lyr3$Ridership.y )
names(finaldataset_rwshd_lyr3)[57]<-"LYRidership"
names(finaldataset_rwshd_lyr3)[3]<-"TYRidership"

finaldataset_rwshd_lyr4 <- merge(finaldataset_rwshd_lyr3,finaldataset_rwshd_lyr3[,c("LYDate","LYRidership")],by.x = "X2YDate",by.y = "LYDate",all.x = TRUE)
finaldataset_rwshd_lyr4 <-finaldataset_rwshd_lyr4[-c(58)]
names(finaldataset_rwshd_lyr4)[57]<-"LYRidership"
names(finaldataset_rwshd_lyr4)[58]<-"X2YRidership"

#add lag 7 days
finaldataset_rwshd_lyr4$LWDate <- finaldataset_rwshd_lyr4$Date -7
finaldataset_rwshd_lyr5 <- merge(finaldataset_rwshd_lyr4,finaldataset_rwshd_lyr4[,c("Date","TYRidership")],by.x = "LWDate",by.y = "Date",all.x = TRUE)
names(finaldataset_rwshd_lyr5)[5]<-"TYRidership"
names(finaldataset_rwshd_lyr5)[60]<-"LWRidership"
finaldataset_rwshd_lyr5[is.na(finaldataset_rwshd_lyr5$LWRidership),"LWRidership"] <- mean(finaldataset_rwshd_lyr5[is.na(finaldataset_rwshd_lyr5$LWRidership),"TYRidership"])

# add lag LY_mines1W_Date
finaldataset_rwshd_lyr5$LY_mines1W_Date <- finaldataset_rwshd_lyr4$LYDate -7
finaldataset_rwshd_lyr6 <-merge(finaldataset_rwshd_lyr5,finaldataset_rwshd_lyr5[,c("LYDate","LYRidership")],by.x = "LY_mines1W_Date",by.y = "LYDate",all.x = TRUE)
names(finaldataset_rwshd_lyr6)[59]<-"LYRidership"
names(finaldataset_rwshd_lyr6)[62]<-"LY_mines1W_Ridership"
finaldataset_rwshd_lyr6[is.na(finaldataset_rwshd_lyr6$LY_mines1W_Ridership),"LY_mines1W_Ridership"] <- mean(finaldataset_rwshd_lyr6[is.na(finaldataset_rwshd_lyr6$LYRidership),"TYRidership"])

# add lag 1 day
finaldataset_rwshd_lyr6$LDDate <- finaldataset_rwshd_lyr6$Date -1
finaldataset_rwshd_lyr6 <- merge(finaldataset_rwshd_lyr6,finaldataset_rwshd_lyr6[,c("Date","TYRidership")],by.x = "LDDate",by.y = "Date",all.x = TRUE)
#head(finaldataset_rwshd_lyr6)
names(finaldataset_rwshd_lyr6)[7]<-"TYRidership"
names(finaldataset_rwshd_lyr6)[64]<-"LDRidership"
finaldataset_rwshd_lyr6[is.na(finaldataset_rwshd_lyr6$LDRidership),"LDRidership"] <- mean(finaldataset_rwshd_lyr6[is.na(finaldataset_rwshd_lyr6$LDRidership),"TYRidership"])

#add lag 2 day
finaldataset_rwshd_lyr6$L2DDate <- finaldataset_rwshd_lyr6$Date -2
finaldataset_rwshd_lyr6 <- merge(finaldataset_rwshd_lyr6,finaldataset_rwshd_lyr6[,c("Date","TYRidership")],by.x = "L2DDate",by.y = "Date",all.x = TRUE)
#head(finaldataset_rwshd_lyr6)
names(finaldataset_rwshd_lyr6)[8]<-"TYRidership"
names(finaldataset_rwshd_lyr6)[66]<-"L2DRidership"
finaldataset_rwshd_lyr6[is.na(finaldataset_rwshd_lyr6$L2DRidership),"L2DRidership"] <- mean(finaldataset_rwshd_lyr6[is.na(finaldataset_rwshd_lyr6$L2DRidership),"TYRidership"])


```

```{r,warning=FALSE}
finaldataset_rwshd_lyr <- finaldataset_rwshd_lyr6[finaldataset_rwshd_lyr6$Date >= '2013-12-31',]

# Use Null value to replace NA values
finaldataset_rwshd_lyr[is.na(finaldataset_rwshd_lyr$LYRidership),"LYRidership"] <- mean(finaldataset_rwshd_lyr[is.na(finaldataset_rwshd_lyr$LYRidership),"TYRidership"])

finaldataset_rwshd_lyr[is.na(finaldataset_rwshd_lyr$X2YRidership),"X2YRidership"] <- mean(finaldataset_rwshd_lyr[is.na(finaldataset_rwshd_lyr$X2YRidership),"TYRidership"])

finaldataset_rwshd_lyr[is.na(finaldataset_rwshd_lyr$LDRidership),"LDRidership"] <- mean(finaldataset_rwshd_lyr[is.na(finaldataset_rwshd_lyr$LDRidership),"TYRidership"])

# Removing Thanksgiving  
finaldataset_rwshd_lyr[finaldataset_rwshd_lyr$TYRidership == 0,"TYRidership"] <- mean(finaldataset_rwshd_lyr[finaldataset_rwshd_lyr$Holiday == "Thanksgiving","TYRidership"])
```

```{r,warning=FALSE}

# Deal with 1/6/2014 - 1/7/2014 school closed ridership fluctuation
finaldataset_rwshd_lyr[finaldataset_rwshd_lyr$X2YDate == '2014-01-06',"X2YRidership"] <- mean(finaldataset_rwshd_lyr[finaldataset_rwshd_lyr$Date == '2014-01-06',"LYRidership"])

finaldataset_rwshd_lyr[finaldataset_rwshd_lyr$LYDate == '2014-01-06',"LYRidership"] <- mean(finaldataset_rwshd_lyr[finaldataset_rwshd_lyr$LYDate == '2014-01-06',"TYRidership"])

finaldataset_rwshd_lyr[finaldataset_rwshd_lyr$Date == '2014-01-06',"TYRidership"] <- mean(finaldataset_rwshd_lyr[finaldataset_rwshd_lyr$Date == '2014-01-06',"LYRidership"])

finaldataset_rwshd_lyr[finaldataset_rwshd_lyr$X2YDate == '2014-01-07',"X2YRidership"] <- mean(finaldataset_rwshd_lyr[finaldataset_rwshd_lyr$Date == '2014-01-07',"LYRidership"])

finaldataset_rwshd_lyr[finaldataset_rwshd_lyr$LYDate == '2014-01-07',"LYRidership"] <- mean(finaldataset_rwshd_lyr[finaldataset_rwshd_lyr$LYDate == '2014-01-07',"TYRidership"])

finaldataset_rwshd_lyr[finaldataset_rwshd_lyr$Date == '2014-01-07',"TYRidership"] <- mean(finaldataset_rwshd_lyr[finaldataset_rwshd_lyr$Date == '2014-01-07',"LYRidership"])

finaldataset_rwshd_lyr[finaldataset_rwshd_lyr$Date == '2014-01-13',"LWRidership"] <- mean(finaldataset_rwshd_lyr[finaldataset_rwshd_lyr$Date == '2014-01-13',"TYRidership"])

```

# We derive month, week of month and quarter of the year variables from ridership date

```{r,warning=FALSE}
finaldataset_rwshd_lyr$month <- as.factor(month(finaldataset_rwshd_lyr$Date))
finaldataset_rwshd_lyr$quarter <- as.factor(quarter(finaldataset_rwshd_lyr$Date))
finaldataset_rwshd_lyr$weekofmonth <- as.factor(ceiling(day(finaldataset_rwshd_lyr$Date)/ 7))
finaldataset_rwshd_lyr$weekdayorend <- as.factor(ifelse(weekdays(finaldataset_rwshd_lyr$Date) %in% c("Saturday", "Sunday"), "weekend", "weekday"))
```

```{r,warning=FALSE}
finaldataset_rwshd_lyr[finaldataset_rwshd_lyr$Date == as.Date('2016-12-31'),][,13:58]<-'OFF'
finaldataset_rwshd_lyr[finaldataset_rwshd_lyr$Date == as.Date('2017-01-01'),][,13:58]<-'OFF'

```

```{r,warning=FALSE}

fixed_LYridership <- read.csv('Easter_Flag.csv')
finaldataset_rwshd_lyr$spw_flag <- fixed_LYridership$Easter_Week_Flag
#finaldataset_rwshd_lyr$spw_flag <- as.factor(finaldataset_rwshd_lyr$spw_flag)

#finaldataset_rwshd_lyr$Chris_flag<- fixed_LYridership$Chris_flag
#finaldataset_rwshd_lyr$Chris_flag<-as.factor(fixed_LYridership$Chris_flag)
#head(finaldataset_rwshd_lyr)
```

# Holidays from current year and previous year are not accurately mapped due to equivalent day's not being holiday's and vice versa. We make sure that they are revised and mapped correctly.

```{r,warning=FALSE}
main_data <- finaldataset_rwshd_lyr
main_data <- subset(main_data, select = c("LYDate","Date","TYRidership","Holiday","day","LYRidership","X2YRidership"))
main_data$Date <- as.Date(main_data$Date)
##updating 2015 new year as a holiday
holiday <- read.csv("updated_calendar_holidays.csv")
holiday$Date <- as.Date(mdy(holiday$Date))
#dates_ly <- main_data %>% filter(Holiday != "NULL") %>% select(LYDate)
new_data_hols <- left_join(main_data,holiday, by = "Date")
new_data_hols$LYHolidayDate <- as.Date(mdy(new_data_hols$LYHolidayDate))
new_data_hols$X2YHolidayDate <- as.Date(mdy(new_data_hols$X2YHolidayDate))
#main_data_date <- main_data$Date
#new_data_sub <- subset(new_data_hols, select = c("Date","TYRidership"))
ridership2012<-read.csv("2012Ridership.csv")
ridership2012$Date <-as.Date(ridership2012$Date)
ridership2012 <- ridership2012%>%group_by(Date) %>%summarise(Ridership = sum(Total.Uses))
usage_pred <- read.csv("Usage.csv")
usage_pred <- subset(usage_pred, select = c("Date","Ridership"))
usage_pred$Date <- as.Date(mdy(usage_pred$Date))
final_ridership <- rbind(ridership2012,usage_pred)
new_data_hols2 <- merge(new_data_hols, final_ridership, by.x = "LYHolidayDate", by.y = "Date", all.x = TRUE)
colnames(new_data_hols2)[13] <- "last_year_holiday_ridership" 
new_data_hols3 <- merge(new_data_hols2, final_ridership, by.x = "X2YHolidayDate", by.y = "Date", all.x = TRUE)
colnames(new_data_hols3)[14] <- "last_two_year_holiday_ridership" 
new_data_hols3_cleaned <- subset(new_data_hols3, select = c("Date","LYDate.x","X2YDate",
                                                            "Holiday.x","LYHolidayDate",
                              "X2YHolidayDate","TYRidership","LYRidership","X2YRidership", 
                              "last_year_holiday_ridership",
                              "last_two_year_holiday_ridership","day"))
## replacing NAs with the mean
mean_1 <- new_data_hols3_cleaned %>% filter(Holiday.x == "New Year's Eve") %>% 
  mutate(avg_ridership = mean(TYRidership))
mean_2 <- new_data_hols3_cleaned %>% filter(Holiday.x == "Suburban School Patrol Day 1") %>% 
  mutate(avg_ridership = mean(TYRidership))
new_data_hols3_cleaned$last_two_year_holiday_ridership[1] <- 47522.75
new_data_hols3_cleaned$last_two_year_holiday_ridership[9] <- 18732.5
#final_data <- if_else(new_data_hols3_cleaned$Holiday.x != "NULL",new_data_hols3_cleaned$LYRidership == new_data_hols3_cleaned$last_year_holiday_ridership,new_data_hols3_cleaned$LYRidership)
final_data <- new_data_hols3_cleaned
final_data$LYRidership[1:100] <- final_data$last_year_holiday_ridership[1:100]
final_data$X2YRidership[1:100] <- final_data$last_two_year_holiday_ridership[1:100]
holiday_revised_ridership <- subset(final_data, select = c("Date","LYRidership","X2YRidership"))
colnames(holiday_revised_ridership) <- c("Date","revised_LY_ridership","revised_x2y_ridership")
```

# Apart from the existing lag days of ridership, we add 14,21,28,35,42 and 49 day lags of ridership.

```{r,warning=FALSE}
finaldataset_rwshd_lyr <- merge(finaldataset_rwshd_lyr,holiday_revised_ridership,all.x = TRUE)
# add lag 14 day
finaldataset_rwshd_lyr$L2wDate <- finaldataset_rwshd_lyr$Date -14
# add lag 21 day
finaldataset_rwshd_lyr$L3wDate <- finaldataset_rwshd_lyr$Date -21
# add lag 28 day
finaldataset_rwshd_lyr$L4wDate <- finaldataset_rwshd_lyr$Date -28
# add lag 35 day
finaldataset_rwshd_lyr$L5wDate <- finaldataset_rwshd_lyr$Date -35
# add lag 42 day
finaldataset_rwshd_lyr$L6wDate <- finaldataset_rwshd_lyr$Date - 42
# add lag 49 day
finaldataset_rwshd_lyr$L7wDate <- finaldataset_rwshd_lyr$Date - 49


finaldataset_rwshd_lyr1 <- merge(finaldataset_rwshd_lyr,final_ridership,by.x = "L2wDate",by.y = "Date",all.x = TRUE)
finaldataset_rwshd_lyr2 <- merge(finaldataset_rwshd_lyr1,final_ridership,by.x = "L3wDate",by.y = "Date",all.x = TRUE)
finaldataset_rwshd_lyr3 <- merge(finaldataset_rwshd_lyr2,final_ridership,by.x = "L4wDate",by.y = "Date",all.x = TRUE)



colnames(finaldataset_rwshd_lyr3)[80:82] <- c("l2w_ridership","l3w_ridership","l4w_ridership")
finaldataset_rwshd_lyr <- finaldataset_rwshd_lyr3
```

# School names were not in format that can be accepted by R. We rename them to comply with R format.

```{r,warning=FALSE}
colnames(finaldataset_rwshd_lyr)[16:61] <-c('Status..Anoka.Hennepin.....Dist.11',
'Status..Blake.Schools..Hopkins..Wayzata..Mpls.',
'Status..Bloomington.....Dist.271',
'Status..Brooklyn.Center.....Dist.286',
'Status..Burnsville.....Dist.191',
'Status..Centennial.....Dist.12',
'Status..Chaska.Area...Dist.112',
'Status..Columbia.Heights.....Dist.13',
'Status..Eden.Prairie.....Dist.272',
'Status..Edina.....Dist.273',
'Status..Elk.River.....Dist.728',
'Status..Farmington',
'Status..Forest.Lake.....Dist.831',
'Status..Fridley.....Dist.14',
'Status..Hastings.....Dist.200',
'Status..Hopkins.....Dist.270',
'Status..Inver.Grove.Heights.....Dist.199',
'Status..Jordan.....Dist.717',
'Status..Lakeville...Dist.194',
'Status..Mahtomedi.....Dist.832',
'Status..Minneapolis.....Dist.1',
'Status..Minnehaha.Academy..Mpls..Blmgtn.',
'Status..Minnetonka.....Dist.276',
'Status..Monticello.....Dist.882',
'Status..Mounds.View.....Dist.621',
'Status..N..St..Paul.Maplewood.Oakdale.Dist.622',
'Status..Orono.....Dist.278',
'Status..Osseo.....Dist.279',
'Status..Prior.Lake.....Dist.719',
'Status..Richfield.....Dist.280',
'Status..Rosemount.Apple.Valley.Eagan...Dist.196',
'Status..Roseville.....Dist.623',
'Status..Shakopee.....Dist.720',
'Status..South.St..Paul.....Dist.6',
'Status..South.Washington.County.....Dist.833',
'Status..Spring.Lake.Park.....Dist.16',
'Status..St.Anthony.New.Brighton.....Dist.282',
'Status..St..Cloud...Dist.742',
'Status..St..Francis.....Dist.15',
'Status..St..Paul.....Dist.625',
'Status..Stillwater',
'Status..Waconia.....Dist.110',
'Status..Wayzata.Plymouth.....Dist.284',
'Status..West.St..Paul.....Dist.197',
'Status..Westonka..Mound......Dist.277',
'Status..White.Bear.Lake.....Dist.624')
write.csv(finaldataset_rwshd_lyr,"finaldataset_rwshd_lyr")
```

# Building the SVM static model by using the MLR package in R. We train a static model with C = 3 and kernel = 'Laplace'. We then predict using the trained model equation. The output of the model will appear as RMSE, MAE and MAPE.

```{r,warning=FALSE}
set.seed(6281)
test <- finaldataset_rwshd_lyr[finaldataset_rwshd_lyr$Date <= '2016-09-30',]
validation <- finaldataset_rwshd_lyr[finaldataset_rwshd_lyr$Date > '2016-09-30',][1:365,]
svmlearner <- makeLearner("regr.ksvm",predict.type = "response")
trainnew <- subset(test,select = -c(1,2,3,4,5,6,7,8,9,10,12,13,14,15,64,65,66,68,69,70,77,78,79))
testnew <- subset(validation,select = -c(1,2,3,4,5,6,7,8,9,10,12,13,14,15,64,65,66,68,69,70,77,78,79))
trainTask <- makeRegrTask(data = trainnew,target = "TYRidership")#7,8,9,10,61
testTask <- makeRegrTask(data = testnew,target = "TYRidership")#,7,8,9,10,61
crossval <- makeResampleDesc("RepCV",folds=10)
```

```{r,warning=FALSE}
optimized <- makeParamSet(
  makeDiscreteParam("kernel",values =c("laplacedot")), #c("vanilladot","polydot","rbfdot","tanhdot","laplacedot")),
  makeIntegerParam("C",lower = 3,upper =3))
tunegrid <- makeTuneControlGrid(resolution = 50)
paramfinal <- tuneParams(learner = svmlearner, resampling = crossval, task = trainTask, par.set = optimized, control = tunegrid, measures = list(mlr::rmse,mlr::mae,mlr::mape),show.info = TRUE)
```

# Prediction for static model

```{r,warning=FALSE}
svmregfinal <- setHyperPars(svmlearner, par.vals = paramfinal$x)
l = mlr::train(svmregfinal,trainTask)
prediction <- predict(l,task = testTask)
actual <- predict(l,task = trainTask)
mlr::performance(prediction,list(mlr::rmse,mlr::mae,mlr::mape))
dfpredict <- cbind(data.frame(prediction$data$truth),data.frame(prediction$data$response),data.frame(validation$Date))
dfpredict <- arrange(dfpredict,prediction.data.truth)

#ggplot() + geom_line(data = dfpredict,aes(x=validation.Date,y=prediction.data.response,color = "red")) + geom_line(data = dfpredict,aes(x=validation.Date,y=prediction.data.truth,color = "green")) +labs(title="Actual Vs Predicted SVM Static",x="Date", y = "Ridership") + scale_color_discrete(name = "Legend",labels = c("Prediction", "Actual"))

dfpredict$error<-abs(dfpredict$prediction.data.truth - dfpredict$prediction.data.response)
```


# Building the SVM 2-week dynamic model by using the MLR package in R. We train the model with C = 3 and kernel = 'Laplace'. We then predict using the trained model equation. The output of the model will appear as RMSE, MAE and MAPE.

```{r,warning=FALSE}
truth1 <- data.frame()
response1 <- data.frame ()

#head(finaldataset_rwshd_lyr)

# Function that returns Root Mean Squared Error
rmse <- function(error)
{
    sqrt(mean(error^2))
}

# Function that returns Mean Absolute Error
mae <- function(error)
{
    mean(abs(error))
}

# Function that returns MAPE
mape <- function(truth, response)
{
    mean(abs((truth-response)/truth))
}
train <- finaldataset_rwshd_lyr[finaldataset_rwshd_lyr$Date <= '2016-09-30',]
  svmlearner <- makeLearner("regr.ksvm",predict.type = "response")
  trainnew = subset(train,select = -c(1,2,3,4,5,6,7,8,9,10,64,65,66,68,69,77,78,79))
  trainTask <- makeRegrTask(data = trainnew,target = "TYRidership") #7,8,9,10,61#,7,8,9,10,61
  crossval <- makeResampleDesc("RepCV",folds=10)
  
  optimized <- makeParamSet(
  makeDiscreteParam("kernel",values =c("laplacedot")), #c("vanilladot","polydot","rbfdot","tanhdot","laplacedot")),
  makeIntegerParam("C",lower = 3,upper =3))
  tunegrid <- makeTuneControlGrid(resolution = 50)
  paramfinal <- tuneParams(learner = svmlearner, resampling = crossval, task = trainTask, par.set = optimized, control = tunegrid, measures = list(mlr::rmse,mlr::mae,mlr::mape),show.info = TRUE)

  svmregfinal <- setHyperPars(svmlearner, par.vals = paramfinal$x)
  l = mlr::train(svmregfinal,trainTask)
  
dynamic_truth <- list()
dynamic_response <- list()
dynamic_error <-list()
for (week in 1:26){
  
  test <- finaldataset_rwshd_lyr[((finaldataset_rwshd_lyr$Date > (as.Date('2016-09-30')-14+week*14))) & (finaldataset_rwshd_lyr$Date <=  (as.Date('2016-09-30')-14+week*28)),]
   
  for_test_append<-finaldataset_rwshd_lyr[finaldataset_rwshd_lyr$Date > '2016-09-30',]

  test<-rbind(test,for_test_append)
  testnew = subset(test,select = -c(1,2,3,4,5,6,7,8,9,10,64,65,66,68,69,77,78,79))
  testTask <- makeRegrTask(data = testnew,target = "TYRidership")
  pred <- predict(l,task = testTask)
  actual <- predict(l,task = trainTask)
  
  truth <- data.frame(pred$data$truth)[1:14,]
  response <-data.frame(pred$data$response)[1:14,]
  
  error <-data.frame(pred$data$truth)-data.frame(pred$data$response)
  error<-error[1:14,]
  dynamic_error <-c(dynamic_error, error)
  dynamic_truth <- c(dynamic_truth, truth)
  dynamic_response <- c(dynamic_response, response)
  
  
  truth1 <- append(truth1,truth)
  response1 <- append(response1,response)
  }
#mean(unlist(dynamic_rmse))
#mean(unlist(dynamic_mae))
#mean(unlist(dynamic_mape))
#train[,-c(1,2,3,4,5,6,60)]
#test[,-c(1,2,3,4,5,6,7,61)]
rmse(unlist(dynamic_error))
mae(unlist(dynamic_error))
mapedataframe <- NULL
mapedataframe$truth <- unlist(dynamic_truth)
mapedataframe$response <- unlist(dynamic_response)
mapedataframe$percerror <- abs(mapedataframe$truth - mapedataframe$response)/mapedataframe$truth
mean(mapedataframe$percerror)

truth1 <- do.call(rbind,truth1)
response1 <- do.call(rbind,response1)
#date1 <- list(ymd(for_test_append$Date))
date1 <- as.data.frame(for_test_append$Date[1:364])

dfpredict <- cbind(date1,truth1,response1)
colnames(dfpredict) <- c("date","actual","prediction")

# ggplot() + geom_line(data = dfpredict,aes(x=date,y=prediction,color = "red")) + 
#  geom_line(data = dfpredict,aes(x=date,y= actual,color = "green")) +
# labs(title="Actual Vs Predicted SVM Dynamic (two weeks)",x="Date", y = "Ridership") + scale_color_discrete(name = "Legend",labels = c("Prediction", "Actual"))




```

# Building the SVM 1-day dynamic model by using the MLR package in R. We train the model with C = 3 and kernel = 'Laplace'. We then predict using the trained model equation. The output of the model will appear as RMSE, MAE and MAPE.

```{r,warning=FALSE}
# Function that returns Root Mean Squared Error
truth1 <- data.frame()
response1 <- data.frame ()

rmse <- function(error)
{
    sqrt(mean(error^2))
}

# Function that returns Mean Absolute Error
mae <- function(error)
{
    mean(abs(error))
}

# Function that returns MAPE
mape <- function(truth, response)
{
    mean(abs((truth-response)/truth))
}

train <- finaldataset_rwshd_lyr[finaldataset_rwshd_lyr$Date <= '2016-09-30',]
  svmlearner <- makeLearner("regr.ksvm",predict.type = "response")
  trainnew = subset(train,select = -c(1,2,3,4,5,6,7,8,9,10,64,65,77,78,79))
  trainTask <- makeRegrTask(data = trainnew,target = "TYRidership") #7,8,9,10,61#,7,8,9,10,61
  crossval <- makeResampleDesc("RepCV",folds=10)
  
  optimized <- makeParamSet(
  makeDiscreteParam("kernel",values =c("laplacedot")), #c("vanilladot","polydot","rbfdot","tanhdot","laplacedot")),
  makeIntegerParam("C",lower = 3,upper =3))
  tunegrid <- makeTuneControlGrid(resolution = 50)
  paramfinal <- tuneParams(learner = svmlearner, resampling = crossval, task = trainTask, par.set = optimized, control = tunegrid, measures = list(mlr::rmse,mlr::mae,mlr::mape),show.info = TRUE)

  svmregfinal <- setHyperPars(svmlearner, par.vals = paramfinal$x)
  l = mlr::train(svmregfinal,trainTask)
  
dynamic_truth <- list()
dynamic_response <- list()
dynamic_error <-list()

for (week in 1:365){
  train <- finaldataset_rwshd_lyr[finaldataset_rwshd_lyr$Date <= (as.Date('2016-09-30')-1+week*1),]
  test <- finaldataset_rwshd_lyr[((finaldataset_rwshd_lyr$Date > (as.Date('2016-09-30')-1+week*1))) & (finaldataset_rwshd_lyr$Date <=  (as.Date('2016-09-30')-14+week*28)),]
  for_test_append<-finaldataset_rwshd_lyr[finaldataset_rwshd_lyr$Date > '2016-09-30',]

  test<-rbind(test,for_test_append)
  testnew = subset(test,select = -c(1,2,3,4,5,6,7,8,9,10,64,65,77,78,79))
  testTask <- makeRegrTask(data = testnew,target = "TYRidership")
  pred <- predict(l,task = testTask)
  actual <- predict(l,task = trainTask)
  
  truth <- data.frame(pred$data$truth)[1,]
  response <-data.frame(pred$data$response)[1,]
  
  error <-data.frame(pred$data$truth)-data.frame(pred$data$response)
  error<-error[1,]
  dynamic_error <-c(dynamic_error, error)
  dynamic_truth <- c(dynamic_truth, truth)
  dynamic_response <- c(dynamic_response, response)
  
   truth1 <- append(truth1,truth)
  response1 <- append(response1,response)
}
#mean(unlist(dynamic_rmse))
#mean(unlist(dynamic_mae))
#mean(unlist(dynamic_mape))
rmse(unlist(dynamic_error))
mae(unlist(dynamic_error))
mapedataframe <- NULL
mapedataframe$truth <- unlist(dynamic_truth)
mapedataframe$response <- unlist(dynamic_response)
mapedataframe$percerror <- abs(mapedataframe$truth - mapedataframe$response)/mapedataframe$truth
mean(mapedataframe$percerror)



truth1 <- do.call(rbind,truth1)
response1 <- do.call(rbind,response1)
#date1 <- list(ymd(for_test_append$Date))
date1 <- as.data.frame(for_test_append$Date[1:365])

dfpredict <- cbind(date1,truth1,response1)
colnames(dfpredict) <- c("date","actual","prediction")

# ggplot() + geom_line(data = dfpredict,aes(x=date,y=prediction,color = "red")) + 
#  geom_line(data = dfpredict,aes(x=date,y= actual,color = "green")) +
# labs(title="Actual Vs Predicted SVM Dynamic (One Day)",x="Date", y = "Ridership") + scale_color_discrete(name = "Legend",labels = c("Prediction", "Actual"))
```
